---
title: "Aides Drug"
author: "Robert Solis!"
date: "6/14/23"
output: html_document
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
triUmp <-read.csv(file="triUmpsal.csv",header=TRUE)
```


# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

A new  preventative aides drug (triUmpsal) is under test in clinical trials. Consider only the DrugUsers, does this drug seem effective?  Hint - use dplyr to obtain that subpopulation with the filter command.

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->
I need to us the filter command to get the data out of trUmph to get it down into a smaller amount that be easier to read.

```{r}
DRUGUSR<-filter(triUmp,DrugUser=="yes")
```



# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

```{r}
DRUGUSR<- DRUGUSR %>%
mutate(across(where(is.character), as.factor))
```

## Descriptive Results

```{r}
# Assuming "infected" and "treatment" are column names in the data frame "Charlie"

```


### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->

```{r}
dd2 <- DRUGUSR %>% group_by(infected, treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```
The first plot shows how many people got infected, it appears around less than 100 people with aids, the non sided people did not get it in percentage terms. 70% of the aids group had taken the drugs and the rest had taken the placebo. The group who took the placebo got infected while the quarter that took the the drug had prevented aids.So that shows the drugs had worked and prevented people from getting aids.

```{r}
# Bar chart for "infected" variable
ggplot(Charlie, aes(x = infected, fill = treatment)) +
  geom_bar(position = "dodge", color = "black") +
  labs(x = "Infected", y = "Count") +
  ggtitle("Bar Chart of Infected and Treatment") +
  scale_fill_manual(values = c("drug" = "blue", "placebo" = "red")) +
  theme(legend.position = "bottom")


```
```{r}
# Bar chart for "treatment" variable
ggplot(Charlie, aes(x = treatment, fill = treatment)) +
  geom_bar(position = "dodge", color = "black") +
  labs(x = "Treatment", y = "Count") +
  ggtitle("Bar Chart of Treatment") +
  scale_fill_manual(values = c("drug" = "blue", "placebo" = "red")) +
  theme(legend.position = "bottom")

```


### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
table1 <- xtabs(~infected + treatment, data=DRUGUSR)
rowPerc(table1)
colPerc(table1)
```

People who end up getting aids about 25% of them have taken the drugs and the other 75% took the placebo. The rest that didnt get anything because of the drug measured 47.59. Thus showing the drug is effective.

## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->



# Discussion and Conclusion

```{r}
library(ggplot2)

# Assuming "infected" and "treatment" are column names in the data frame "Charlie"

# Bar chart for "infected" variable
ggplot(Charlie, aes(x = infected, fill = treatment)) +
  geom_bar(position = "dodge", color = "black") +
  labs(x = "Infected", y = "Count") +
  ggtitle("Bar Chart of Infected and Treatment") +
  scale_fill_manual(values = c("drug" = "blue", "placebo" = "red")) +
  theme(legend.position = "bottom")

# Bar chart for "treatment" variable
ggplot(Charlie, aes(x = treatment, fill = treatment)) +
  geom_bar(position = "dodge", color = "black") +
  labs(x = "Treatment", y = "Count") +
  ggtitle("Bar Chart of Treatment") +
  scale_fill_manual(values = c("drug" = "blue", "placebo" = "red")) +
  theme(legend.position = "bottom")

```


<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variables causes the other, or whether they perhaps are related for some other reason.  Watch the chapter 6 videos from the George Town videos collection.!-->
